---
title: "HW1"
output: html_notebook
author: Evan Batzer
date: 1/14/2019
---
__1 - Computation__

Compute the median annual spending for each agency.

unzip, unz, list.files, lapply, tapply, median, log, hist, file.info,
system.time

```{r}
library(tidyverse); library(data.table)
zip_file_path = "./data/awards.zip"
files = unzip(zip_file_path, list = TRUE)

fnames = files$Name

fnames = fnames[-1]

sumSpending = function(fnames){
  
  output <- list()
  processsummary <- list()
  counter = 1
  
  for(fname in fnames){
    
    #d = fread(paste('unzip -cq ', zip_file_path, fname), 
    #          blank.lines.skip = TRUE)
    f = unz(zip_file_path, fname)
    d = read.csv(f)
     
    d$date = as.Date(d$period_of_performance_start_date)
    d$year = year(d$date)
    
    testthat::expect_true(length(unique(d$funding_agency_name)) == 1)
    total =  tapply(d$total_obligation, INDEX = d$year, FUN = sum)
    
    output[[counter]] = data.frame(agency = unique(d$funding_agency_name),
                                   total,
                                   year = names(total))
    
    processsummary[[counter]] = data.frame(agency = unique(d$funding_agency_name),
                                           filename = fname,
                                           nrows = nrow(d))
    
    counter = counter + 1
  }
  
  return(list(medians = do.call(rbind, output),
              process = do.call(rbind, processsummary)))
}

output = sumSpending(fnames)

View(output[[2]])
```

1. Which agencies have the highest median annual spending?

```{r, warning=FALSE}
median_df = output[[1]] %>%
  group_by(agency) %>%
  summarise(mspend = median(total))

head(median_df %>% arrange(desc(mspend)))
```

```{r, warning=FALSE}
output = sumSpending(fnames)

median_df = output[[1]] %>%
  group_by(agency) %>%
  summarise(mspend = median(total))

head(median_df %>% arrange(desc(mspend)))
```


2. Qualitatively describe the distribution of median annual spending.

```{r}
med_plot = median_df %>% 
  ggplot(aes(x = mspend)) +
  geom_histogram(fill = "lightblue", color = "black") +
  ggtitle("Median Spending Distribution")

med_plot
```

3. Qualitatively describe the distribution of the logarithm of the median annual spending.
Plot the histogram.

```{r}
logmed_plot = median_df %>% 
  ggplot(aes(x = log(mspend))) +
  geom_histogram(fill = "lightblue", color = "black") +
  ggtitle("Log Median Spending Distribution")

logmed_plot
```

Plotting both together for writeup

```{r}
gridExtra::grid.arrange(med_plot, logmed_plot, nrow = 1)
```

4. Is there a clear separation between agencies that spend a large amount of money, and
those which spend less money?


1. Qualitatively describe the distribution of the file sizes.

```{r}
spendfiles = unzip(zip_file_path, list = TRUE)
spendfiles %>% ggplot(aes(x = Length)) +
  geom_histogram(fill = "lightblue", color = "black") +
  ggtitle("Distribution of File Length")  
  
```

2. How does the size of the file relate to the number of rows in that file?

```{r}
inner_join(spendfiles, output[[2]], by = c("Name" = "filename")) %>%
  ggplot(aes(x = Length,
             y = nrows)) +
  geom_point(size = 4, fill = "lightblue", color = "black", pch = 21) +
  stat_smooth(se = FALSE, method = "lm")
```

3. How long does it take to process all the data?

```{r}
system.time(sumSpending(fnames))
```

4. Do you think this same approach you took work for 10 times as many files? What if
each file was 10 times larger?

5. How do you imagine you could make it faster?



